What makes musicians infer teaching intentions?

### 2023-01-07 paper/manuscript.Rmd
correct some grammatical errors
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

If performers' intentions are understood by learners with basic musical skills, they should be able to distinguish teaching recordings from performing recordings. Moreover, the same modulations of pedagogical performance parameters observed in our previous experiments such as slower demonstration and exaggerated performance, should be used to infer teaching intentions by the learners. In Experiment 1 participants listened to performances that followed simple musical scales. In Experiment 2 participants listened to performances of a more complex piece.

@@

## Stimuli
We selected stimuli from the performances obtained in our previous experiments [@tominaga_2022]. Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive technique of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from "A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment. The stimuli were performed around 80 quarter-beats per minute.

@@

However, dynamics (KV) seems continued to be used as cue to teaching intentions despite the increased musical complexity. Particularly performances with larger contrasts between forte and piano seem to be considered as teaching. Moreover, performances with both exaggerated forte and piano were considered to be for teaching. Thus, loudness (dynamics) might be used as a reliable cue to infer teaching intentions regardless of the complexity of a musical piece.

# General discussion

@@

To examine which features of piano performance make musicians infer pedagogical intentions, we performed correlation and multiple regression analysis. Across the two experiments, it was found that loudness, particularly larger contrast between forte and piano, strongly contributed to participants’ judgments when listening to dynamics recordings. On the other hand, unlike what we predicted, slower performance was only considered to be for teaching for both techniques in Experiment 1 where the stimuli consisted of simple musical scales. Also, we could not find that performances with exaggerated articulation (i.e., longer legato and shorter staccato) were more likely to be judged as teaching in Experiment 2. This indicates that some characteristics of performance are not necessarily used or reliable to infer performers’ intentions when listening to musically complex pieces. Certain modulations may have been too subtle to be noticed by the participants. This would be in line with @gabrielsson_2010 who reported that tempo and loudness reign over other musical factors and seem to be easier for listeners to judge. Another possibility is that a wide range of modulations is considered as expressive in more complex pieces. This raises the possibility that it is easier to learn and teach musical expressive techniques when musical complexity is low.

The piece used in Experiment 2 was from actual sonatinas. Although sonatina pieces are generally aimed to be simple and easy for beginner pianists to play, they add a multitude of musical properties to a performance that is not present in musical scales. Not only do performers produce their communicative intentions such as emotion, but composers can express their feelings by handling various musical notations such as melody, rhythm and harmony [@gabrielsson_2010]. In our study, most of the staccato parts of the piece used in Experiment 2 (*Figure \ref{fig:stim2}*, (2)) were comprised of the same repetitive notes. This implied that they could be only played in a staccato style. Dynamics may be more independent of the structure of the piece compared to articulation. Therefore, participants might have had difficulties in recognising teaching intentions from articulation in Experiment 2. This would imply that the ease of difficulty with which teaching intentions can be derived vary across different pieces depend on a variety of musical properties.

>>>>

If performers' intentions are understood by learners with basic musical skills, they should be able to distinguish teaching recordings from performing recordings. Moreover, the same modulations of pedagogical performance parameters observed in our previous experiments such as slower demonstration and exaggerated performance, should be used to infer teaching intentions by the learners. In Experiment 1, participants listened to performances that followed simple musical scales. In Experiment 2, participants listened to performances of a more complex piece.

@@

## Stimuli
We selected stimuli from the performances obtained in our previous experiments [@tominaga_2022]. Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive technique of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from "A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment. The stimuli were performed at around 80 quarter-beats per minute.

@@

However, dynamics (KV) seems to continue to be used as a cue to teaching intentions despite the increased musical complexity. Particularly performances with larger contrasts between forte and piano seem to be considered as teaching. Moreover, performances with both exaggerated forte and piano were considered to be for teaching. Thus, loudness (dynamics) might be used as a reliable cue to infer teaching intentions regardless of the complexity of a musical piece.

# General Discussion

@@

To examine which features of piano performance make musicians infer pedagogical intentions, we performed correlation and multiple regression analysis. Across the two experiments, it was found that loudness, particularly larger contrasts between forte and piano, strongly contributed to participants’ judgments when listening to dynamics recordings. On the other hand, unlike what we predicted, slower performance was only considered to be for teaching for both techniques in Experiment 1 where the stimuli consisted of simple musical scales. Also, we could not find that performances with exaggerated articulation (i.e., longer legato and shorter staccato) were more likely to be judged as teaching in Experiment 2. This indicates that some characteristics of performance are not necessarily used or reliable to infer performers’ intentions when listening to musically complex pieces. Certain modulations may have been too subtle to be noticed by the participants. This would be in line with @gabrielsson_2010 who reported that tempo and loudness reign over other musical factors and seem to be easier for listeners to judge. Another possibility is that a wide range of modulations is considered as expressive in more complex pieces. This raises the possibility that it is easier to learn and teach musical expressive techniques when musical complexity is low.

The piece used in Experiment 2 was from actual sonatinas. Although sonatina pieces are generally aimed to be simple and easy for beginner pianists to play, they add a multitude of musical properties to a performance that is not present in musical scales. Not only do performers produce their communicative intentions such as emotion, but composers can express their feelings by handling various musical notations such as melody, rhythm and harmony [@gabrielsson_2010]. In our study, most of the staccato parts of the piece used in Experiment 2 (*Figure \ref{fig:stim2}*, (2)) were comprised of the same repetitive notes. This implied that they could be only played in a staccato style. Dynamics may be more independent of the structure of the piece compared to articulation. Therefore, participants might have had difficulties in recognising teaching intentions from articulation in Experiment 2. This would imply that the ease or difficulty with which teaching intentions can be derived varies across different pieces depending on a variety of musical properties.

### 2023-01-07 paper/manuscript.tex
correct some grammatical errors
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

@@

\abstract{
Perceiving pedagogical intentions is vital when learning skills from others. Here we investigated whether the modulations that expert pianists produce when they intend to teach are also perceived by listeners as conveying pedagogical intentions. Musicians listened to piano recordings where a musical expressive technique of either articulation or dynamics was implemented. They were asked to judge whether each recording was produced for teaching purposes or not. The findings in Experiment 1 with a simple musical scale demonstrated that the accuracy of musicians' judgments was significantly above chance. Also, slower tempo contributed to musicians' judgments as teaching regardless of the techniques. Moreover, performances with exaggeration for each technique were more likely to be judged as teaching. Experiment 2 aimed to replicate the findings of Experiment 1 with a more complex piece of music. The accuracy of musicians' judgments was significantly above chance. In contrast to Experiment 1, only exaggerated dynamics were more likely to be judged as teaching. Taken together, the results indicate that recognition of some teaching intentions may be facilitated by low musical complexity.
}

@@

We selected stimuli from the performances obtained in our previous experiments (Tominaga et al., 2022). Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive technique of either articulation (\emph{Figure \ref{fig:stim1}}, A) or dynamics (\emph{Figure \ref{fig:stim1}}, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from ``A Dozen a Day - Play with Ease in Many Keys'' by Edna-Mae Burnam and modified for the experiment. The stimuli were performed around 80 quarter-beats per minute.

@@

However, dynamics (KV) seems continued to be used as cue to teaching intentions despite the increased musical complexity. Particularly performances with larger contrasts between forte and piano seem to be considered as teaching. Moreover, performances with both exaggerated forte and piano were considered to be for teaching. Thus, loudness (dynamics) might be used as a reliable cue to infer teaching intentions regardless of the complexity of a musical piece.

\hypertarget{general-discussion}{%
\section{General discussion}\label{general-discussion}}

@@

To examine which features of piano performance make musicians infer pedagogical intentions, we performed correlation and multiple regression analysis. Across the two experiments, it was found that loudness, particularly larger contrast between forte and piano, strongly contributed to participants' judgments when listening to dynamics recordings. On the other hand, unlike what we predicted, slower performance was only considered to be for teaching for both techniques in Experiment 1 where the stimuli consisted of simple musical scales. Also, we could not find that performances with exaggerated articulation (i.e., longer legato and shorter staccato) were more likely to be judged as teaching in Experiment 2. This indicates that some characteristics of performance are not necessarily used or reliable to infer performers' intentions when listening to musically complex pieces. Certain modulations may have been too subtle to be noticed by the participants. This would be in line with Gabrielsson and Lindström (2010) who reported that tempo and loudness reign over other musical factors and seem to be easier for listeners to judge. Another possibility is that a wide range of modulations is considered as expressive in more complex pieces. This raises the possibility that it is easier to learn and teach musical expressive techniques when musical complexity is low.

The piece used in Experiment 2 was from actual sonatinas. Although sonatina pieces are generally aimed to be simple and easy for beginner pianists to play, they add a multitude of musical properties to a performance that is not present in musical scales. Not only do performers produce their communicative intentions such as emotion, but composers can express their feelings by handling various musical notations such as melody, rhythm and harmony (Gabrielsson \& Lindström, 2010). In our study, most of the staccato parts of the piece used in Experiment 2 (\emph{Figure \ref{fig:stim2}}, (2)) were comprised of the same repetitive notes. This implied that they could be only played in a staccato style. Dynamics may be more independent of the structure of the piece compared to articulation. Therefore, participants might have had difficulties in recognising teaching intentions from articulation in Experiment 2. This would imply that the ease of difficulty with which teaching intentions can be derived vary across different pieces depend on a variety of musical properties.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tominaga_2022a}{}}%
Tominaga, A., Knoblich, G., \& Sebanz, N. (2022). \emph{The sound of teaching music: {Expert} pianists' performance modulations for novices}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/wzuyj}

>>>>

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

@@

\abstract{%
Perceiving pedagogical intentions is vital when learning skills from others. Here we investigated whether the modulations that expert pianists produce when they intend to teach are also perceived by listeners as conveying pedagogical intentions. Musicians listened to piano recordings where a musical expressive technique of either articulation or dynamics was implemented. They were asked to judge whether each recording was produced for teaching purposes or not. The findings in Experiment 1 with a simple musical scale demonstrated that the accuracy of musicians' judgments was significantly above chance. Also, slower tempo contributed to musicians' judgments as teaching regardless of the techniques. Moreover, performances with exaggeration for each technique were more likely to be judged as teaching. Experiment 2 aimed to replicate the findings of Experiment 1 with a more complex piece of music. The accuracy of musicians' judgments was significantly above chance. In contrast to Experiment 1, only exaggerated dynamics were more likely to be judged as teaching. Taken together, the results indicate that recognition of some teaching intentions may be facilitated by low musical complexity.
}

@@

We selected stimuli from the performances obtained in our previous experiments (Tominaga et al., 2022). Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive technique of either articulation (\emph{Figure \ref{fig:stim1}}, A) or dynamics (\emph{Figure \ref{fig:stim1}}, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from ``A Dozen a Day - Play with Ease in Many Keys'' by Edna-Mae Burnam and modified for the experiment. The stimuli were performed at around 80 quarter-beats per minute.

@@

However, dynamics (KV) seems to continue to be used as a cue to teaching intentions despite the increased musical complexity. Particularly performances with larger contrasts between forte and piano seem to be considered as teaching. Moreover, performances with both exaggerated forte and piano were considered to be for teaching. Thus, loudness (dynamics) might be used as a reliable cue to infer teaching intentions regardless of the complexity of a musical piece.

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

@@

To examine which features of piano performance make musicians infer pedagogical intentions, we performed correlation and multiple regression analysis. Across the two experiments, it was found that loudness, particularly larger contrasts between forte and piano, strongly contributed to participants' judgments when listening to dynamics recordings. On the other hand, unlike what we predicted, slower performance was only considered to be for teaching for both techniques in Experiment 1 where the stimuli consisted of simple musical scales. Also, we could not find that performances with exaggerated articulation (i.e., longer legato and shorter staccato) were more likely to be judged as teaching in Experiment 2. This indicates that some characteristics of performance are not necessarily used or reliable to infer performers' intentions when listening to musically complex pieces. Certain modulations may have been too subtle to be noticed by the participants. This would be in line with Gabrielsson and Lindström (2010) who reported that tempo and loudness reign over other musical factors and seem to be easier for listeners to judge. Another possibility is that a wide range of modulations is considered as expressive in more complex pieces. This raises the possibility that it is easier to learn and teach musical expressive techniques when musical complexity is low.

The piece used in Experiment 2 was from actual sonatinas. Although sonatina pieces are generally aimed to be simple and easy for beginner pianists to play, they add a multitude of musical properties to a performance that is not present in musical scales. Not only do performers produce their communicative intentions such as emotion, but composers can express their feelings by handling various musical notations such as melody, rhythm and harmony (Gabrielsson \& Lindström, 2010). In our study, most of the staccato parts of the piece used in Experiment 2 (\emph{Figure \ref{fig:stim2}}, (2)) were comprised of the same repetitive notes. This implied that they could be only played in a staccato style. Dynamics may be more independent of the structure of the piece compared to articulation. Therefore, participants might have had difficulties in recognising teaching intentions from articulation in Experiment 2. This would imply that the ease or difficulty with which teaching intentions can be derived varies across different pieces depending on a variety of musical properties.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tominaga_2022}{}}%
Tominaga, A., Knoblich, G., \& Sebanz, N. (2022). Expert pianists make specific exaggerations for teaching. \emph{Scientific Reports}, \emph{12}(1), 21296. \url{https://doi.org/10.1038/s41598-022-25711-3}
